Writing and Reading Empirical Papers

### 2019-10-10 static/slides/paper-slides.Rmd
Add midterm review questions, fix paper project slides

1. Introduction
2. Literature Review
3. Theory/Model 
4. Data Description
5. Empirical Model
6. Results/Implications
7. Bibliography

@@

- Get to your research question ASAP!  Make it the first sentence even.
- Hook your reader
    - Who cares?  Why is this important?  Why is this relevant? How does this affect people?
		- Statistics and background information can often help

@@

.content-box-green[
.green[**Example**]: I estimate the relationship between paper length and grades by using a simple OLS regression using sample data collected from previous classes. I find that there is a strong positive effect, that students who write longer papers earn higher grades. On average, for every additional page written, grades improve by 2 points.
]

@@

- These are economics papers, so you must describe some **economic theory** behind the question you are asking and answering
- Most scholarly papers have a formal economic model, which then generates predictions that they test for with data
- You do not need a model, but you do need to discuss economic principles or concepts that are relevant
    - Often there may be multiple theories that might conflict, or our expectations might not be clear (these are the best papers!)
    - There may be a significant tradeoff between competing goals, values, or expectations

@@

- Rule I: Show your data! Show us basic summary statistics and any patterns
    - your judgment: we don't want or need to see *everything* -- what do you think is interesting or important?
    - Plots $>$ Tables $>$ Nothing
- Good ideas to *always* have:
    1. A table of variables and their description
    2. A table of summary statistics of variables
    3. Plots of most important variables and interactions (histograms, boxplots, scatterplots, etc)

@@

```{r, fig.retina=3, fig.align="center", out.width="60%"}
papers<-papers %>%
  mutate(Sex=as.factor(Sex),
         Sex=recode(Sex, `F` = "Female", M = "Male")) %>%
  mutate(Female = ifelse(Sex=="Female",1,0),
         Morning = ifelse(Time=="Morning",1,0),
         Hood=ifelse(School=="Hood",1,0),
         Econometrics=ifelse(Class=="Econometrics",1,0))

@@

.pull-left[
```{r}
papers %>%
  count(Morning) %>%
  knitr::kable(., format="html")
```

@@

# Data: Counts of Categorical Variables II

@@

```{r, fig.retina=3, fig.align="center", out.width="60%"}
pages_hist<-ggplot(data = papers)+
  aes(x = Pages)+
  geom_histogram(color="white", fill="indianred",breaks=seq(0,24,2))+
  scale_x_continuous(breaks=seq(0,24,2))+
  scale_y_continuous(breaks=seq(0,50,10))+
  labs(x = "Number of Pages Written",
       y = "Number of Papers")+
    theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
pages_hist
```

@@

```{r, fig.retina=3, fig.align="center", out.width="60%"}
pages_hist+facet_wrap(~Sex)
```

@@

```{r, fig.retina=3, fig.align="center", out.width="60%"}

@@

- Are your results **robust** across different model specifications?
    - Do the size(s) of the marginal effect(s) you care about change or reverse direction? Become/lose significance?
- At minimum, you must run several models, including a multivariate regression
    - Run several variations of your model with and without controls (e.g. just $Y$ and $X$, $Y$ and $X_1$ and $X_2$, etc.)
    - Check for nonlinearities: polynomials, logs, etc.

@@

.center[
![:scale 90%](https://www.dropbox.com/s/qtd2uwvr52jvkcu/drwhocontext.jpg?raw=1)
]

>>>>

1. Introduction

2. Literature Review

3. Theory/Model

4. Data Description

5. Empirical Model

6. Results/Implications

7. Bibliography

@@

- Get to your research question ASAP!  Make it the first sentence even.

- Hook your reader
    - Who cares?  Why is this important?  Why is this relevant? How does this affect people?
		- Statistics and background information can often help

@@

.content-box-green[
.green[**Example**]: I estimate the relationship between paper length and grades by using a simple OLS regression using sample data collected from previous classes. I find that there is a strong positive effect, that students who write longer papers earn higher grades. On average, for every additional page written, grades improve by less than a point. These results are robust to a number of different model specifications and controls.
]

@@

- These are economics papers, so you must describe some **economic theory** behind the question you are asking and answering

- Most scholarly papers have a formal economic model, which then generates predictions that they test for with data

- You do not need a model, but you do need to discuss economic principles or concepts that are relevant
    - Often there may be multiple theories that might conflict, or our expectations might not be clear (these are the best papers!)
    - There may be a significant tradeoff between competing goals, values, or expectations

@@

- .shout[Show your data!] Show us basic summary statistics and any patterns
    - Use your judgment: .onfire[we don't want or need to see *everything*]
    - What do you think is *interesting* or *important*?
    - Plots $>$ Tables $>$ Words $>$ Nothing

- Good ideas to *always* have:
    1. A table(s) of variables and their description
    2. A table(s) of summary statistics of variables
    3. A table of correlations of key variables
    3. Plots of most important variables and interactions (histograms, boxplots, scatterplots, etc)

@@

```{r, fig.retina=3, fig.align="center", out.width="100%"}
papers<-papers %>%
  mutate(Sex=as.factor(Sex),
         Sex=recode(Sex, `F` = "Female", M = "Male")) %>%
  mutate(Female = ifelse(Sex=="Female",1,0),
         Morning = ifelse(Time=="Morning",1,0),
         Hood=ifelse(School=="Hood",1,0),
         Econometrics=ifelse(Class=="Econometrics",1,0))

@@

.pull-left[
```{r}
papers %>%
  count(Time) %>%
  knitr::kable(., format="html")
```

@@

@@

```{r, fig.retina=3, fig.align="center", out.width="50%"}
pages_hist<-ggplot(data = papers)+
  aes(x = Pages)+
  geom_histogram(color="white", fill="indianred",breaks=seq(0,24,2))+
  scale_x_continuous(breaks=seq(0,24,2))+
  scale_y_continuous(breaks=seq(0,50,10))+
  labs(x = "Number of Pages Written",
       y = "Number of Papers")+
    theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
pages_hist
```

@@

```{r, fig.retina=3, fig.align="center", out.width="50%"}
pages_hist+facet_wrap(~Sex)
```

@@

```{r, fig.retina=3, fig.align="center", out.width="50%"}

@@

- Are your results **robust** across different model specifications?
    - Do the size(s) of the marginal effect(s) you care about change or reverse direction? Become/lose significance?

- At minimum, you must run several models, including a multivariate regression
    - Run several variations of your model with and without controls (e.g. just $Y$ and $X$, $Y$ and $X_1$ and $X_2$, etc.)
    - Check for nonlinearities: polynomials, logs, etc.

@@

.center[
![:scale 45%](https://www.dropbox.com/s/qtd2uwvr52jvkcu/drwhocontext.jpg?raw=1)
]
